The 2- and 3-point surface correlation functions calculations: from novel exact continuous approach to improving methodology for discrete images

### 2023-05-11 paper.tex

\section{Computation of surface functions for digital images}
\label{sec:algo}
Here we briefly describe the algorithm for computation of surface-surface,
surface-void and surface-surface-surface correlation functions for digital
images of porous media \cite{Samarin}. This discrete algorithm can calculate surface
correlation functions for all possible directions and correlation lengths. It
consists of two steps: edge detection and computation of autocorrelation. Next,
we present some improvements in this approach that later will be verified withthe help of
exact continuous method.

@@

\begin{figure*}
  \centering
  \subfigure[Surface-surface function calculated for a ball with radius
    $R = 0.24$. Image dimensions are  $450 \times 450 \times 450$.]{
    \includegraphics[width=0.3\linewidth]{images/ball-3d-ss.png}
    \label{fig:fss-ball}}
  \hfill
  \subfigure[Surface-void function calculated for a disk with radius
    $R = 0.2$. Image dimensions are $2000 \times 2000$.]{
    \includegraphics[width=0.3\linewidth]{images/fsv-disk.png}
    \label{fig:fsv-disk}}
  \hfill
  \subfigure[Surface-void function calculated for a ball with radius
    $R = 0.2$. Image dimensions are $300 \times 300 \times 300$.]{
    \includegraphics[width=0.3\linewidth]{images/fsv-ball.png}
    \label{fig:fsv-ball}}
  \caption[]{Comparison of filters $H$ and $H'$ for cases in which newly
    developed algorithms \cref{sec:fss-2d}, \cref{sec:fsss-3d} are not
    applicable. Theoretical curves are provided for each case.}
  \label{fig:not-covered}
\end{figure*}
With the new filter the calculation of $F_{ss}$ becomes more precise also for
three-dimensional images. This can be seen in comparison with theoretical values
for known cases like the one of a ball. On \cref{fig:fss-ball} there are plots
of $F_{ss}$ function calculated with filters $H$ and $H'$ along with a
theoretical curve. It can be verified that filter $H'$ is better indeed by
calculating
\begin{equation*}
  error = \sqrt{\frac{\sum_i (F_{ss}^{calc}(x_i) -
      F_{ss}^{theory}(x_i))^2}{\sum_i F_{ss}^{theory}(x_i)^2}}
\end{equation*}
where $x_i$ belong to a set
$(\epsilon, 2R - \epsilon) \cup (2R + \epsilon, 1)$. Taking $\epsilon = 0.01$ we
get $error = 0.006$ for filter $H$ and $error = 0.00017$ for filter
$H'$. Calculation of surface-void function also benefits from the new filter
(\cref{fig:fsv-disk}, \cref{fig:fsv-ball}).

Unfortunately the filter $H'$ has some drawbacks in comparison with $H$. Firstly
$H$ has faster convergence with increase of image resolution. Our
acceptance criterion developed in \textcolor{red}{cite our paper again} works
also for $H'$, but sometimes it's possible to calculate a poor estimate of
$F_{ss}$ using $H$ but not $H'$. Secondly, $H'$ behaves worse that $H$ in points
of discontinuity of $F_{ss}$ as can be seen on \cref{fig:fss-disk}.

>>>>

\section{Computation of surface functions for digital images}
\label{sec:algo}
Here we briefly describe the algorithm for computation of surface-surface,
surface-void and surface-surface-surface correlation functions for digital
images of porous media \cite{Samarin}. This discrete algorithm can calculate
surface correlation functions for all possible directions and correlation
lengths. It consists of two steps: edge detection and computation of
autocorrelation. Next, we present some improvements in this approach that later
will be verified with the help of exact continuous method.

@@

\begin{figure*}
  \centering
  \subfigure[Surface-surface function calculated for a ball with radius
    $R = 0.24$. Image dimensions are  $450 \times 450 \times 450$.]{
    \includegraphics[width=0.3\linewidth]{images/ball-3d-ss.png}
    \label{fig:fss-ball}}
  \hfill
  \subfigure[Surface-void function calculated for a disk with radius
    $R = 0.2$. Image dimensions are $2000 \times 2000$.]{
    \includegraphics[width=0.3\linewidth]{images/fsv-disk.png}
    \label{fig:fsv-disk}}
  \hfill
  \subfigure[Surface-void function calculated for a ball with radius
    $R = 0.2$. Image dimensions are $300 \times 300 \times 300$.]{
    \includegraphics[width=0.3\linewidth]{images/fsv-ball.png}
    \label{fig:fsv-ball}}
  \caption[]{Comparison of filters $H$ and $H'$ for cases in which newly
    developed algorithms \cref{sec:fss-2d}, \cref{sec:fsss-3d} are not
    applicable. Theoretical curves are provided for each case.}
  \label{fig:not-covered}
\end{figure*}
With the new filter the calculation of $F_{ss}$ becomes more precise also for
three-dimensional images. This can be seen in comparison with theoretical values
for known cases like the one of a ball. On \cref{fig:fss-ball} there are plots
of $F_{ss}$ function calculated with filters $H$ and $H'$ along with a
theoretical curve. It can be verified that filter $H'$ is better indeed by
calculating
\begin{equation}
  error = \sqrt{\frac{\sum_i (F_{ss}^{calc}(x_i) -
      F_{ss}^{theory}(x_i))^2}{\sum_i F_{ss}^{theory}(x_i)^2}}
  \label{eq:error}
\end{equation}
where $x_i$ belong to a set $(\epsilon, 2R - \epsilon)$. Taking
$\epsilon = 0.01$ we get $error = 0.05$ for filter $H$ and $error = 0.03$ for
filter $H'$. Calculation of surface-void function also benefits from the new
filter (\cref{fig:fsv-disk}, \cref{fig:fsv-ball}).

Unfortunately the filter $H'$ has some drawbacks in comparison with
$H$. Firstly, $H$ has faster convergence with increase of image resolution. To
demonstrate this we can use an acceptance parameter $C_{0.5}$ developed to
determine if an image is suitable for processing with our algorithm or not. This
parameter is defined as
\begin{equation*}
  C_{0.5} = \frac{\text{Power of the lower half of the image's
      spectrum}}{\text{Power of the image}}
\end{equation*}
In this expression the words ``Power of the image'' are understood as $L^2$ norm
of the image and ``the lower half of the \dots spectrum'' is the part of the
spectrum (Fourier image) with frequencies $[-\omega_N/2, \omega_N/2]$ where
$\omega_N$ is a Nyquist frequency for the image. It's important to firstly
subtract the zeroth frequency (mean value) from the image before calculation of
$C_{0.5}$. The image is suitable for processing if the acceptance criterion
$C_{0.5} > 0.97$ holds. This criterion works for both filters $H$ and $H'$, but
$H$ results in smaller error when $C_{0.5}$ is slightly less than 0.97. For
example, consider images of a disk used for calculation of $F_{ss}$ on
\cref{fig:fss-disk}. On \cref{fig:error} we provide values of $C_{0.5}$ for
those images and relative errors in $F_{ss}$ (\cref{eq:error}) for both
filters. It can be seen that $H$ gives twice the lower error than $H'$ when used
for an image with $C_{0.5} \approx 0.82$.

Secondly, $H'$ behaves worse that $H$ in points
of discontinuity of $F_{ss}$ as can be seen on \cref{fig:fss-disk}.

\begin{figure*}[!hpt]
  \centering
  \subfigure[$C_{0.5}$ criterion]{
    \includegraphics[width=0.45\linewidth]{images/c05.png}
    \label{fig:c05}}
  \hfill
  \subfigure[Relative errors in computation of $F_{ss}$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-errors.png}
    \label{fig:relerr}}
  \caption[]{Relative errors in function $F_{ss}$ for the case of a disk
    (\cref{fig:fss-disk}) for filters $H$ and $H'$ and the acceptance parameter
    $C_{0.5}$.}
  \label{fig:error}
\end{figure*}
